---
title: "Comparison with Christensen paper"
author: "Renata Diaz"
date: "12/13/2018"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Side-by-side comparison of LDATS results (and Portal rodents vignette) with analysis from Christensen et al 2018. 

## LDATS Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

```{r, eval = T}
library(LDATS)
```

## Christensen 2018 analysis

Download Christensen 2018 analysis scripts & data files from [Extreme-events-LDA repo:](https://github.com/emchristensen/Extreme-events-LDA)

Main Analysis Scripts:

  * rodent_LDA_analysis.R main script for analyzing rodent community change using LDA
  
  * rodent_data_for_LDA.R contains a function that creates the rodent data table used in analyses
  
  * AIC_model_selection.R contains functions for calculating AIC for different candidate LDA models
  
  * changepointmodel.r contains change-point model code
  
  * LDA-distance.R function for computing Hellinger distance analyses
    
Data:

  * Rodent_table_dat.csv table of rodent data, created by rodent_data_for_LDA.R

Figure scripts:

  * LDA_figure_scripts.R contains functions for making main plots in manuscript (Fig 1). Called from rodent_LDA_analysis.R


```{r download Christensen, eval = FALSE}

files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')

for(i in 1:length(files_to_download)) {
  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
                destfile = paste0('christensen-ecology-files/', files_to_download[i]))
}

rm(files_to_download)
rm(i)

```


## Data

The Portal rodents control data is included in the LDATS package:

```{r LDATS data}

data(rodents)

ldats_dat <- rodents[[1]]
ldats_dates <- rodents[[2]]
head(ldats_dat)

```

Load the data used in Christensen et al: 

```{r Ch data}
source('christensen-ecology-files/rodent_data_for_LDA.r')
dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = F)

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% rownames(dat)) %>% select(period,date)
dates = period_dates$date

ch_dat <- dat

ch_dates <- dates

rm(list = c('dat', 'dates', 'period_dates', 'moondat',
            'create_rodent_table'))

```


Compare paper data to LDATS data:

```{r rodent data comparison} 

compare <- ldats_dat == ch_dat

compare_rows <- vector(length = nrow(compare)) 

for(i in 1:nrow(compare)) {
  if(sum(compare[i, ]) == 21) {
    compare_rows[i] <- TRUE
  } else {
    compare_rows[i] <- FALSE
  }
}

length(which(compare_rows == F))
ldats_dat[ which(compare_rows == F), ]
ch_dat[ which(compare_rows == F), ]
```


There are 16 rows where the data included in LDATS differs from the paper data. This looks like it is because the LDATS data is not adjusted to account for trapping effort, but the paper data has divided all census counts by the actual number of plots trapped and multiplied by 8 to account for incompletely-trapped censuses. 

Double-check:

```{r Ch rodent adjustment, eval = F}

source('christensen-ecology-files/rodent_data_for_LDA.r')
diagnose_dat = create_rodent_table(period_first = 1,
                          period_last = 436,
                          selected_plots = c(2,4,8,11,12,14,17,22),
                          selected_species = c('BA','DM','DO','DS','NA','OL','OT','PB','PE','PF','PH','PI','PL','PM','PP','RF','RM','RO','SF','SH','SO'), diagnose = T)

adjusted_table <- diagnose_dat[[1]]
raw_table <- diagnose_dat[[2]]
nplots <- diagnose_dat[[3]]
nplots <- nplots[ which(nplots$period <= 436), ]

compare_raw <- raw_table == ldats_dat
length(which(compare_raw == F))

which(compare_rows == F) 
which(nplots$x < 8)

```

The lines that are different are the lines that were adjusted; if you don't adjust the paper data you get the same table as the LDATS data. 

```{r data adjustment cleanup, include = FALSE}
rm(list = c('diagnose_dat', 'nplots', 'compare', 'compare_raw', 'adjusted_table', 'i', 'compare_rows', 'raw_table', 'create_rodent_table'))
```


For now I will use the *adjusted rodent table* because this will avoid artificially low abundances for incompletely trapped periods.

```{r switch to adjusted rodent data}

ldats_dat <- ch_dat
rodents$document_term_table <- ch_dat

```

