---
title: "Latent Dirichlet Allocation Time Series (LDATS)"
author: "Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

## Overview 

This vignette details the code base for the **LDATS** package. It was
constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Analytical Pipeline

The full LDATS analysis can be executed using the `LDA_TS` function,
which is the top-level of the coding pipeline's hierarchy of
functions:

* `LDA_TS`
  * `check_LDA_TS_inputs`
    * `check_timename()`
    * `check_formulas`
    * `check_nchangepoints`
    * `check_weights`
    * `check_control`
    * `check_document_term_table`
    * `check_document_covariate_table`	
    * `check_topics`
    * `check_seeds`
  * `LDA_set`
    * `LDA`
  * `select_LDA`
  * `TS_on_LDA`
    * `expand_TS`
    * `TS`
      * `est_changepoints`
      * `est_regressors`
      * `summarize_TS`
  * `select_TS`
  * `package_LDA_TS`

## Controls Lists

`LDA_TS_controls_list`